\chapter{Newtonian Gravity}
In Newtonian physics we assume that we have absolute space and time that can be described by the a set of numbers $x^1,x^2,x^3,t$.
We express the coordinates as functions of time.
\section{Forces}
The force $\vec{F}_{AB}$, which a massive body $A$ with mass $m_A$ exerts on another massive body $B$ with mass $m_B$, is given by
\begin{equation}
    \vec{F}_{AB}=-m_B\frac{G\textsubscript{N}m_A}{r^2}\vec{e}_r\, ,
\end{equation}
where $G\textsubscript{N}$ denotes \emph{Newton's constant}, numerically equal
to $G\textsubscript{N}\approx \unitfrac[6.673\cdot 10^{-11}]{m^3}{kg\,s}$.
Although there is no need for $G\textsubscript{N}$ to be constant over time,
there is evidence that the relative variation is less than $10^{-12}$ per year.
The force can be expressed in terms of \emph{gravitational potential} $\Phi$:
\begin{equation}
    \vec{F}_{AB}=-m_B\nabla\left(-\frac{G\textsubscript{N}m_A}{r}\right)=:
    -m_B\nabla\Phi(\vec{r}_B)\, .
\end{equation}
Given $N$ particles labeled by $n$, the total force $B$ experiences is
\begin{equation}
    \vec{F}_{B}=-\sum_n \vec{F}_{nB}\, .
\end{equation}
The potential at $\vec{r}$ is then easily found to be
\begin{equation}
    \Phi(\vec{r})=-G\textsubscript{N}\sum_n\frac{m_n}{|\vec{r}-\vec{r}_n|}\, .
\end{equation}
In general, we assume a mass distribution $\varrho(\vec{r})$ and the sum is
replaced by an integral:
\begin{equation}
    \Phi(\vec{r}) = -G\textsubscript{N}\int_{\Reals^3}\dif{\vec{r}^{\prime}}
    \frac{\varrho(\vec{r}^{\prime})}{|\vec{r}-\vec{r}^{\prime}|}\,.
\end{equation}
\section{Comparison with electrostatics}
The classical theory of gravity bears a striking similarity to electrostatics.
To make this clearer, we introduce the gravitational field
$\vec{g}(\vec{r}):= -\nabla\Phi(\vec{r})$.
\begin{table}[htb]
    \caption{Comparison of electrostatics and Newtonian gravity.}
    \begin{center}
        \begin{tabulars}{lll}
            \toprule
            &Newtonian Gravity&Electrostatics\\
            \midrule
            Force&$\displaystyle\vec{F}=q\frac{kQ}{r^2}\vec{e}_r$&$\vec{F}=m\frac{G\textsubscript{N}M}{r^2}\vec{e}_r$\\
            Potential
            &$\Phi\textsubscript{el}(\vec{r})=q\frac{kQ}{r}$
            &$\Phi\textsubscript{g}(\vec{r})=m\frac{G\textsubscript{N}M}{r}$\\
            Field
            &$\vec{E}(\vec{r})=-\nabla\Phi\textsubscript{el}(\vec{r})$
            &$\vec{g}(\vec{r})=-\nabla\Phi\textsubscript{g}(\vec{r})$\\
            Laplace Equation
            &$\Delta\Phi\textsubscript{el}=-4\pi k\varrho\textsubscript{el}(\vec{r})$&
            $\Delta\Phi\textsubscript{g}=4\pi
            G\textsubscript{N}\varrho\textsubscript{g}(\vec{r})$
            \\
            \bottomrule
        \end{tabulars}
    \end{center}
\end{table}
\begin{example}[Field of a spherical mass distribution]
Assume we have a spherical mass distribution, i.e.\ $\varrho(\vec{r})=\varrho(r)$.
By symmetry considerations it follows, that the gravitational field can be
expressed by
\begin{equation}
    \vec{g}(\vec{r})=g(r)\vec{e}_r\, .
\end{equation}
We integrate the divergence of the field over a ball $B$ of radius $r$
\begin{equation}
    \int_B\dif{\vec{r}}\,\nabla\vec{g}=-\int_B\dif{\vec{r}}\, \Delta\Phi
    = -4\pi G\textsubscript{N}\int_B\dif{\vec{r}}\,\varrho(r)= -4\pi
    G\textsubscript{N} M\, ,
\end{equation}
where $M$ is the mass enclosed in $B$. On the other hand we can use Gauss's theorem to deduce
\begin{equation}
    \int_B\dif{\vec{r}}\,\nabla\vec{g}=\oint_{\partial B}\dif{\vec{A}}\cdot
    \vec{g} = \oint_{\Omega}\dif{\Omega}\, g(r)r^2=4\pi r^2g(r)\, .
\end{equation}
Together the gravitational field is given by
\begin{equation}
    \vec{g}(r)=-\frac{G\textsubscript{N}M}{r^2}\vec{e}_{r}\, .
\end{equation}
\end{example}
% \subsection{Inertial systems}
% \begin{definition}
% An inertial system is a system in which force-free particles move with constant uniform velocity on straight lines.
% \end{definition}
\subsection*{Weak Equivalence Principle (WEP)}
Newtons first law reads
\begin{equation}
    \vec{F}=m\textsubscript{I} \ddot{\vec{x}} \, ,
\end{equation}
where $m\textsubscript{I}$ is the inertial mass that works against the acceleration of the body.
The force which a body with ``active'' mass $m\textsubscript{g,act.}$ exerts on another
body with mass $m\textsubscript{g,pass.}$ is given by 
\begin{equation}
    \vec{F}=m\textsubscript{g,pass.}\frac{G\textsubscript{N}m\textsubscript{g,act.}}{r^2}\vec{e}_r \, .
\end{equation}
A priori, there is no reason to assume any relation between this masses.
The first question one might ask is whether the active and the passive mass are equal.
Suppose we have two masses $A$ and $B$. Using Newtons first law, we can explicitly write
\begin{equation}
    m^{B}_{\text{i}}\ddot{\vec{x}}=\vec{F}_{AB}=-	m^{B}_{\text{p}}\frac{G\textsubscript{N}m^{A}_{\text{a}}}{r^2}\vec{e}_r \, .
\end{equation}
%TODO finish chapter 1